Gradient Descent for Numerical MAX-CSP

### 2022-04-30 proj/paper/paper.tex
change opt

\documentclass[a4paper]{article}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{dsfont}
\usepackage{paralist}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyvrb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{paralist}
\usepackage[svgname]{xcolor}
\usepackage{enumerate}
\usepackage{array}
\usepackage{times}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{comment}
\usepackage{environ}
\usepackage{times}
\usepackage{textcomp}
\usepackage{caption}
\usepackage[square,numbers]{natbib}

@@

\setlength\parindent{0pt} % Removes all indentation from paragraphs

@@

\title{Stochastic Gradient Descent for Numerical MAX-CSP} % Title
\author{
Derek Paulsen \\
}

@@

\section{Abstract}
\section{Introduction}
% TODO
\section{Background}

\subsubsection{Constraint satisfaction problems}

Constraint satisfaction problems (CSPs) encompasses many different problems. Formally a 
constraint satisfaction problem (CSP) is defined as a triple $\langle X, D, C \rangle$, where 
% XXX this is from wikipedia https://en.wikipedia.org/wiki/Constraint_satisfaction_problem
\begin{align*}
	X &= \{X_1, ..., X_n\} \text{ the set of variables}\\
	D &= \{D_1, ..., D_n\} \text{ the domains of each respective variable}\\
	C &= \{C_1, ..., C_m\} \text{ the set of constraints}\\
\end{align*}

A classic example of CSP, is the 3-SAT problem. In this case
the boolean variables that appear in at least one clause would be $X$. The domain for 
each variable would be $D_i = \{true, false\}$ and the constraints would all be 
three variable disjunctive clauses. The 3-SAT problem of course known to be NP-Complete.
In this paper we focus on a subset of CSPs called Numerical MAX-CSP. 
In typical a CSP all constraints must be satisfied, however in MAX-CSP the goal is simply to 
satisfy as may constraints as possible, additionally the domains are the variables 
are numerical (e.g. the real numbers).

While this problem has certainly been considered before, we were only able to find one previous 
work that directly address this issue. 
%TODO Add citation here

@@

\begin{align*}
\min_{x,z}\quad &1^Tz\\
s.t. \quad Ax - \epsilon z &\leq 0\\
		w &\in \mathds{R}^n\\
		z &\in \{0,1\}^m\\
\end{align*}

@@

\subsubsection{Gradient Descent}

%TODO rewrite this

@@

@@

$$
f(x) = \text{HardSigmoid}(Aw - \gamma)
$$
Where, 
$$
x\in R, \quad \text{HardSigmoid}(x) = \begin{cases} 0 \text{ if } x \leq 0\\ 1 \text{ if } x \geq 1\\ x \text{ otherwise}\end{cases}
$$

For gradient updates we start with $\eta = .5$ which decays exponentially every 25 gradient updates. We do a total 
of 100 gradient updates and then back track to find the $w$ (i.e. the one that satisfies as many constraints as 
possible). We repeat this process starting with random $w$ until we reach a preset time limit.

@@

We ran our experiments on a server with an AMD Ryzen Threadripper 2950X (16c/32t) processor with 
64GB of RAM. Our MAX-CSP instances are generated from real world datasets, with the number of constraints 
ranging from \red{TODO : max num const} to \red{TODO : max num const}. For the baseline solutions we 
leverage Gurobi, which takes advantage of all threads available.

\section{Discussion}
\section{Conclusions}

%\bibliographystyle{abbrvnat}
%\bibliography{references}
\end{document}

>>>>

\documentclass[a4paper]{article}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{booktabs}
\usepackage{dsfont}
\usepackage{paralist}
\usepackage{epstopdf}
\usepackage{tabularx}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{multicol}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyvrb}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{paralist}
\usepackage[svgname]{xcolor}
\usepackage{enumerate}
\usepackage{array}
\usepackage{times}
\usepackage{url}
\usepackage{fancyhdr}
\usepackage{comment}
\usepackage{environ}
\usepackage{times}
\usepackage{textcomp}
\usepackage{caption}
\usepackage[square,numbers]{natbib}

@@

@@

\title{Gradient Descent for Numerical MAX-CSP} % Title
\author{
Derek Paulsen \\
}

@@

\section{Abstract}
\section{Introduction}

% TODO

Many real world problems can be modeled as constraint 
satisfaction problems. The problems range from logic puzzles 
like sudoku to tuning search engines. Briefly, a constraint satisfaction 
problem is simply the task of finding a valid assignment of variables 
which satisfy as set of constraints. Many classic problems in theoretical computer 
science fit this model, such as the NP-Complete 3-SAT and circuit SAT problems. In this paper 
look variant of CSP, called numerical MAX-CSP. In this variant each variable is a real number
and all of the constraints are linear inequalities, the goal then is to satisfy as 
many linear inequalities as possible.

The rest of this paper is then structured as follows. First, we give background on the problem 
and the basis for our solution. Next we motivate solving this problem by tying it to 
tuning a keyword search engine with labeled data. We then move on to describe a baseline 
solution by modeling the problem as mixed integer linear program and discuss the issues with 
this solution. Next, we present our algorithm and compare it to the baseline solution. Finally, 
we discuss the experimental results and give directions for future work.

\section{Background}

\subsection{Constraint satisfaction problems}

Constraint satisfaction problems (CSPs) encompasses many different problems. Formally a 
constraint satisfaction problem (CSP) is defined as a triple $\langle X, D, C \rangle$, where 
% XXX this is from wikipedia https://en.wikipedia.org/wiki/Constraint_satisfaction_problem
\begin{align*}
	X &= \{X_1, ..., X_n\} \text{ the set of variables}\\
	D &= \{D_1, ..., D_n\} \text{ the domains of each respective variable}\\
	C &= \{C_1, ..., C_m\} \text{ the set of constraints}\\
\end{align*}
A classic example of a CSP, is the 3-SAT problem. In this case
the boolean variables that appear in at least one clause would be $X$. The domain for 
each variable would be $D_i = \{true, false\}$ and the constraints would all be 
three variable disjunctive clauses. The 3-SAT problem of course known to be NP-Complete.
In this paper we focus on a subset of CSPs called Numerical MAX-CSP. 
In typical a CSP all constraints must be satisfied, however in MAX-CSP the goal is simply to 
satisfy as may constraints as possible, additionally the domains are the variables 
are numerical (e.g. the real numbers).

While this problem has certainly been considered before, we were only able to find one previous 
work that directly addresses this issue. 
%TODO Add citation here

@@

\begin{align*}
\min_{x,z}\quad &1^Tz\\
s.t. \quad Ax - \epsilon z &\leq b\\
		w &\in \mathds{R}^n\\
		z &\in \{0,1\}^m\\
\end{align*}

@@

\subsection{Gradient Descent}

@@

Where each row, $a$, of $A$ comes corresponds to 
$$
a = (d_{i,y} - d_{i,x}) \odot q_i  \quad i \in \{1,...n\} \quad x,y \in \{1,...,k\}, l_{i,x} = relevant, l_{i,y} = irrelevant
$$

That is, the pairwise difference between the components of a matching and non-matching document, multiplied by the 
relevant query. We note that with this setup, 
$$
a^Tw < 0 \implies (q\odot w)^Td_{i,x} > (q\odot w)^Td_{i,y}
$$
That is, the relevant document will be scored higher than the non-relevant document for the query. Solving this 
MILP then corresponds to minimizing the number of irrelevant documents that need to be 
examined in order to find the relevant document for each query.

@@

$$
f(x) = \text{HardSigmoid}(Aw - \gamma)
$$
Where $ \gamma = -1$ and, 
$$
x\in R, \quad \text{HardSigmoid}(x) = \begin{cases} 0 \text{ if } x \leq 0\\ 1 \text{ if } x \geq 1\\ x \text{ otherwise}\end{cases}
$$

For gradient updates we start with $\eta = .5$ which decays exponentially every 25 gradient updates. We do a total 
of 500 gradient updates and then back track to find the best $w$ (i.e. the one that satisfies as many constraints as 
possible). We then repeat this process starting with random $w$ until we reach a preset time limit.

@@

We ran our experiments on a server with an AMD Ryzen Threadripper 2950X (16c/32t) processor with 
64GB of RAM. Our MAX-CSP instances are generated from real world datasets, with the number of constraints 
ranging from 7769 to 5994778. For the baseline solutions we 
leverage Gurobi, a commercial solver with an academic license. We choose gurobi because it is consistently 
one of the top performing available solvers \cite{benchmark}.

\begin{tabular}{|l|r|r|r|}
\toprule
Dataset &  \# of Constraints &  \# of Columns &  \% Non-Zero \\
\midrule
         $D_0$ &                   7769 &                 14 &         66.24 \\
         $D_1$ &                  19350 &                 18 &         61.46 \\
         $D_2$ &                 111974 &                 18 &         72.83 \\
         $D_3$ &                 196213 &                 30 &         51.62 \\
         $D_4$ &                3711208 &                124 &         25.09 \\
         $D_5$ &                5994778 &                 22 &         68.58 \\
\bottomrule
\end{tabular}

\bibliographystyle{abbrvnat}
\bibliography{references}
\end{document}
